---
title: "hw4"
author: "Group 4"
date: '2022-11-10'
output: html_document
---

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)

```

```{r}
insurance.train <- read.csv('https://raw.githubusercontent.com/djunga/DATA621HW4/main/insurance_train.csv')
insurance.eval <- read.csv('https://raw.githubusercontent.com/djunga/DATA621HW4/main/insurance_eval.csv')
```

```{r}
head(insurance.train)
```

### Removing unwanted chars "Z_" from all columns
```{r}

insurance.train[] <- Map(gsub, "^z_", "", insurance.train)

head(insurance.train)

```

### Identify columns with missing values. Count number of 'na' in columns
```{r}
colSums(is.na(insurance.train))                 # Count missing values
```

### Summarize the data
```{r}
summary(insurance.train)
```

### All the columns are of String type. Convert the relevant one to numeric

```{r}
ins_train <- transform(insurance.train, INDEX = as.numeric(INDEX),
                       TARGET_FLAG = as.numeric(TARGET_FLAG), 
                    TARGET_AMT = as.numeric(TARGET_AMT), KIDSDRIV = as.numeric(KIDSDRIV), AGE= as.numeric(AGE), HOMEKIDS= as.numeric(HOMEKIDS),
                    YOJ = as.numeric(YOJ), TIF = as.numeric(TIF), CLM_FREQ = as.numeric(CLM_FREQ), MVR_PTS = as.numeric(MVR_PTS), CAR_AGE = as.numeric(CAR_AGE) 
                    )

summary(ins_train)
head(ins_train)
```

## Variation in data

### Visualizing distribution

```{r}
ggplot(ins_train) + 
  geom_bar(mapping = aes(x = TARGET_FLAG))
```


### Let's see the distribution based on the vehicle type
```{r}
ggplot(data = ins_train) +
  geom_bar(mapping = aes(x = CAR_TYPE))
```


### Let's see the distribution based on the vehicle's age

```{r}
ggplot(data = ins_train) +
  geom_bar(mapping = aes(x = CAR_AGE))
```

### Let's see the distribution based on the driver's age
```{r}
ggplot(data = ins_train) +
  geom_bar(mapping = aes(x = AGE))
```

### Plotting Age of the driver by Target amount

```{r}
ggplot(data = ins_train, mapping = aes(x = AGE, y = TARGET_AMT)) + 
  geom_point()
```


### Plotting Age of the driver by Target flag

```{r}
ggplot(data = ins_train, mapping = aes(x = AGE, y = TARGET_FLAG)) + 
  geom_point()
```



### I found CAR_AGE to be an interesting attribute. Let's try to dig deeper into that subset

```{r}

ins_train <- ins_train %>% 
  mutate(NEW_CAR = ifelse(CAR_AGE < 6, 1, 0))
```


### Subset NEW_CAR data
```{r}
new_car_dataset <- filter(ins_train, NEW_CAR==1)
old_car_dataset <- filter(ins_train, NEW_CAR==0)

```


### Plotting Age of the driver by Target amount with new cars

```{r}
ggplot(new_car_dataset) + 
  geom_bar(mapping = aes(x = TARGET_FLAG))

ggplot(old_car_dataset) + 
  geom_bar(mapping = aes(x = TARGET_FLAG))


ggplot(data = new_car_dataset, mapping = aes(x = AGE, y = TARGET_AMT)) + 
  geom_point()

ggplot(data = old_car_dataset, mapping = aes(x = AGE, y = TARGET_AMT)) + 
  geom_point()

```

### Please note that the percentage of new car getting into a crash is much greater than an older car having an accident, irrespective of the driver's age 

```{r}

ggplot(data = ins_train, mapping = aes(x = CAR_TYPE, y = TARGET_FLAG)) + 
  geom_point()

```

### Subsetting the dataset with all crash vehicles
```{r}

crash_car_dataset <- filter(ins_train, TARGET_FLAG==1)

ggplot(data = crash_car_dataset, mapping = aes(x =  reorder(CAR_TYPE, TARGET_AMT, FUN = median), y = TARGET_AMT)) +
  geom_boxplot()

```

### Covariation between CarType & Target Amount

```{r}
ggplot(data = crash_car_dataset, mapping = aes(x = TARGET_AMT)) + 
  geom_freqpoly(mapping = aes(colour = CAR_TYPE), binwidth = 10000)
```



